### 2025-04-17 docs/Homeownership_BankOfMomAndDad.tex
Update behind regressions

\begin{table}[htbp]
	\begin{adjustbox}{center}
		\def\sym#1{\ifmmode^{#1}\else\(^{#1}\)\fi}
		\begin{threeparttable}
			\singlespacing
			\caption{Descriptive Statistics (Means), Households Aged 25-44}\label{tab:descrstats}
			\begin{tabular}{l  r@{\hspace{1.\tabcolsep}}r r@{\hspace{1.\tabcolsep}}r r@{\hspace{1.\tabcolsep}}r r@{\hspace{1.\tabcolsep}}r}
				\toprule
				& \multicolumn{2}{c}{All}  &  \multicolumn{2}{c}{Renter}   & \multicolumn{2}{c}{Owner}    \\ \cmidrule(lr){2-3} \cmidrule(lr){4-5} \cmidrule(lr){6-7}
				Receiver 		& \multicolumn{1}{c}{No} & \multicolumn{1}{c}{Yes} & \multicolumn{1}{c}{No} & \multicolumn{1}{c}{Yes} & \multicolumn{1}{c}{No} & \multicolumn{1}{c}{Yes}    \\  \midrule
				\input ../tabfig/descr/descr_trsnf_disc_mean.tex
				\midrule
				Age                    & \multicolumn{2}{c}{25-31} & \multicolumn{2}{c}{32-38}  & \multicolumn{2}{c}{39-44}   \\ \cmidrule(lr){2-3} \cmidrule(lr){4-5} \cmidrule(lr){6-7}
				Receiver       & \multicolumn{1}{c}{No} & \multicolumn{1}{c}{Yes} & \multicolumn{1}{c}{No} & \multicolumn{1}{c}{Yes} & \multicolumn{1}{c}{No} & \multicolumn{1}{c}{Yes}    \\ 
				 \midrule
				\input ../tabfig/descr/descr_trsnf_age_mean.tex
				\midrule
				Wealth Tertile                    & \multicolumn{2}{c}{Tertile 1}  & \multicolumn{2}{c}{Tertile 2}  & \multicolumn{2}{c}{Tertile 3}    \\ \cmidrule(lr){2-3} \cmidrule(lr){4-5} \cmidrule(lr){6-7}
				Receiver       & \multicolumn{1}{c}{No} & \multicolumn{1}{c}{Yes} & \multicolumn{1}{c}{No} & \multicolumn{1}{c}{Yes} & \multicolumn{1}{c}{No} & \multicolumn{1}{c}{Yes}    \\ 
				\midrule
				\input ../tabfig/descr/descr_trsnf_wlt_mean.tex
		\bottomrule
			\end{tabular}
			{\footnotesize
				\textit{Notes:} Data from the PSID Transfer, Individual, and Family modules. Weighted using family weights. Transfer, wealth, and income measured in 1000s of 2016 USD. }
		\end{threeparttable}

@@

It is perhaps not so surprising that the effect of transfer receipt is so big. First, beyond the 2013 transfer supplement (which has the smallest effect, 1.6 pp) I only capture transfers above \$10,000. Standard models, including the quantiative model in this paper, predicts that renters follow a threshold rule: for any renter that chooses to rent, there exists a wealth level at which they prefer to own (Indeed, the regression results also show that a 10\% increase in net worth increases the probability of becoming a homeowner by 4 percentage points.). Hence, large transfers can shift households above the trehsold if they were `close'. Second, these transfers are not exogenous: indeed, it is likely that some of these transfers given exactly because the household is buying.

\begin{table}
	\center
	\begin{threeparttable}
		\caption{The Transition to Ownership}
		\label{tab:newowners_short}
		\small 
				\input{../tabfig/regr/newowners_short.tex}

@@

\subsubsection{Households with Wealthier Parents Buy More Expensive Houses}
I first show that households with wealthier parents buy more expensive homes, controlling for household characteristics.

I run an ordinary least squares (OLS) regression on house purchase values for first-time owners and test whether lagged parental wealth (logged) is positively associated with larger house purchases $(\beta_1)$, controlling for lagged log income, lagged log net worth, education, age, log household size, marriage, state, time, and race:
\begin{equation*}
\ln{HVal_{i}} = \beta_1 \ln(Wealth)_{p(i),t-2} + \beta_2 \ln(Inc_{i,t-2}) + \beta_3 \ln(NetWorth_{i,t-2}) + \mathbf{\gamma}{X_{i,t}} + \varepsilon_{i}.
\end{equation*}
Parents' wealth, as well as household income and net worth, are log-transformed. Households are denoted by $i$ and their parents by $p(i)$. First-time buyers are households who have never been observed as homeowners before.

Column 1 of Table \ref{tab:hypo} reports the results. Households with wealthier parents buy larger houses: A 1\% increase in parental wealth is associated with a 0.049\% increase in the purchase value of the child's house, and the effect is significant at the 5\% level. The effect of parental wealth is about the same as the child's own net worth (0.065\%).

\begin{table}
	\center
	\begin{threeparttable}
		\caption{Housing Choices and Parental Wealth}
		\label{tab:hypo}
		\small 
				\input{../tabfig/regr/hypoII.tex}

		{\begin{footnotesize}\begin{flushleft}
		\textit{Notes:} Standard errors in parentheses. `Behind' refers to whether the household is behind on a mortgage. Wealth, income, parental wealth, mortgage, family size, and house values are logged. All regressions include year and state fixed effects and control for age and age-squared of both the child and parent. Specifications 1-3 use ordinary least squares while specifications 4 and 5 use random and fixed effects, respectively. \textsuperscript{+}$p<0.10$, \textsuperscript{*}$p<0.05$, \textsuperscript{**}$p<0.01$, \textsuperscript{***}$p<0.001$.
		\end{flushleft}\end{footnotesize}}		
	\end{threeparttable}
\end{table}

\subsubsection{Households with Wealthier Parents Are Less Likely to Be Behind on Mortgages} 
To realize the benefits of homeownership, households must be able to maintain their mortgage payments, since failure to do so can lead to distressed and forced sales. I now show that households with wealthier parents are less likely to be behind on their mortgages, even though they buy more expensive houses. The PSID has collected this data since 2009, and 3.9\% of owners aged 25 to 44 are behind, on average, while 9.2\% have been behind at least once.

The sample is limited to the first time a household is observed as owners, and the control variables are the same as in the previous regression:
\begin{equation*}
Behind_i = \beta_1 \ln(Wealth)_{p(i),t-2} + \beta_2 \ln(Inc_{i,t-2}) + \beta_3 \ln(NetWorth_{i,t-2}) + \mathbf{\gamma}{X_{i,t}} + \varepsilon_{i}.
\end{equation*}
In Column 2 of Table \ref{tab:hypo}, the outcome variable is whether households are behind on the mortgage payments the first time they are observed as owners. Parental wealth in the period before purchase decreases the probability that a household will ever be behind: A 1\% increase in parental wealth decreases the probability of being behind by 0.016 percentage points and is significant at the 1\% level. Parental wealth has a larger effect than the child's net worth. In Column 3, the outcome variable is whether first-time homeowners will ever be observed to be behind on their mortgage, and coefficients and significance levels are virtually unchanged. The sample size increases in these regressions, since households who bought before 2009 are now included.

>>>>

\begin{table}[htbp]
	\begin{adjustbox}{center}
		\small
		\def\sym#1{\ifmmode^{#1}\else\(^{#1}\)\fi}
		\begin{threeparttable}
			\singlespacing
			\caption{Descriptive Statistics (Means), Households Aged 25-44}\label{tab:descrstats}
			\begin{tabular}{l  r@{\hspace{1.\tabcolsep}}r r@{\hspace{1.\tabcolsep}}r r@{\hspace{1.\tabcolsep}}r r@{\hspace{1.\tabcolsep}}r}
				\toprule
				& \multicolumn{2}{c}{All}  &  \multicolumn{2}{c}{Renter}   & \multicolumn{2}{c}{Owner}    \\ \cmidrule(lr){2-3} \cmidrule(lr){4-5} \cmidrule(lr){6-7}
				Receiver 		& \multicolumn{1}{c}{No} & \multicolumn{1}{c}{Yes} & \multicolumn{1}{c}{No} & \multicolumn{1}{c}{Yes} & \multicolumn{1}{c}{No} & \multicolumn{1}{c}{Yes}    \\  \midrule
				\input ../tabfig/descr/descr_trsnf_disc_mean.tex
				\midrule
				Age                    & \multicolumn{2}{c}{25-31} & \multicolumn{2}{c}{32-38}  & \multicolumn{2}{c}{39-44}   \\ \cmidrule(lr){2-3} \cmidrule(lr){4-5} \cmidrule(lr){6-7}
				Receiver       & \multicolumn{1}{c}{No} & \multicolumn{1}{c}{Yes} & \multicolumn{1}{c}{No} & \multicolumn{1}{c}{Yes} & \multicolumn{1}{c}{No} & \multicolumn{1}{c}{Yes}    \\ 
				 \midrule
				\input ../tabfig/descr/descr_trsnf_age_mean.tex
				\midrule
				Wealth Tertile                    & \multicolumn{2}{c}{Tertile 1}  & \multicolumn{2}{c}{Tertile 2}  & \multicolumn{2}{c}{Tertile 3}    \\ \cmidrule(lr){2-3} \cmidrule(lr){4-5} \cmidrule(lr){6-7}
				Receiver       & \multicolumn{1}{c}{No} & \multicolumn{1}{c}{Yes} & \multicolumn{1}{c}{No} & \multicolumn{1}{c}{Yes} & \multicolumn{1}{c}{No} & \multicolumn{1}{c}{Yes}    \\ 
				\midrule
				\input ../tabfig/descr/descr_trsnf_wlt_mean.tex
		\bottomrule
			\end{tabular}
			{\footnotesize
				\textit{Notes:} Data from the PSID Transfer, Individual, and Family modules. Weighted using family weights. Transfer, wealth, and income measured in 1000s of 2016 USD. }
		\end{threeparttable}

@@

\textcolor{red}{Missing regression equation}
It is perhaps not so surprising that the effect of transfer receipt is so big. First, beyond the 2013 transfer supplement (which has the smallest effect, 1.6 pp) I only capture transfers above \$10,000. Standard models, including the quantiative model in this paper, predicts that renters follow a threshold rule: for any renter that chooses to rent, there exists a wealth level at which they prefer to own (Indeed, the regression results also show that a 10\% increase in net worth increases the probability of becoming a homeowner by 4 percentage points.). Hence, large transfers can shift households above the trehsold if they were `close'. Second, these transfers are not exogenous: indeed, it is likely that some of these transfers given exactly because the household is buying.

\begin{table}
	\begin{threeparttable}
		\caption{The Transition to Ownership}
		\label{tab:newowners_short}
		\small 
				\input{../tabfig/regr/newowners_short.tex}

@@

\subsubsection{Households with Wealthier Parents Are Less Likely to Be Behind on Mortgages} 
\textcolor{red}{!! [[rewrite this]] !!}
To realize the benefits of homeownership, households must be able to maintain their mortgage payments, since failure to do so can lead to distressed and forced sales. I now show that households with wealthier parents are less likely to be behind on their mortgages, even though they buy more expensive houses. The PSID has collected this data since 2009, and 3.9\% of owners aged 25 to 44 are behind, on average, while 9.2\% have been behind at least once.

The sample is limited to the first time a household is observed as owners, and the control variables are the same as in the previous regression:
\begin{equation*}
Behind_i = \beta_1 \ln(Wealth)_{p(i),t-2} + \beta_2 \ln(Inc_{i,t-2}) + \beta_3 \ln(NetWorth_{i,t-2}) + \mathbf{\gamma}{X_{i,t}} + \varepsilon_{i}.
\end{equation*}
In Column 2 of Table \ref{tab:hypo}, the outcome variable is whether households are behind on the mortgage payments the first time they are observed as owners. Parental wealth in the period before purchase decreases the probability that a household will ever be behind: A 1\% increase in parental wealth decreases the probability of being behind by 0.016 percentage points and is significant at the 1\% level. Parental wealth has a larger effect than the child's net worth. In Column 3, the outcome variable is whether first-time homeowners will ever be observed to be behind on their mortgage, and coefficients and significance levels are virtually unchanged. The sample size increases in these regressions, since households who bought before 2009 are now included.
\begin{table}
	\begin{threeparttable}
		\caption{Housing Choices and Parental Wealth}
		\label{tab:hypo}
		\small 
				\input{../tabfig/regr/hypoII.tex}

		{\begin{footnotesize}\begin{flushleft}
		\textit{Notes:} Standard errors in parentheses. `Behind' refers to whether the household is behind on a mortgage. Wealth, income, parental wealth, mortgage, family size, and house values are logged. All regressions include year and state fixed effects and control for age and age-squared of both the child and parent. Specifications 1 and 2 use OLS while specifications 3 and 4 use RE and FE, respectively. \textsuperscript{+}$p<0.10$, \textsuperscript{*}$p<0.05$, \textsuperscript{**}$p<0.01$, \textsuperscript{***}$p<0.001$.
		\end{flushleft}\end{footnotesize}}		
	\end{threeparttable}
\end{table}
